Fast Sorting on Modern Computers

### 2005-07-24 docs/tech_report/method.tex
updated heap scripts to remove 16 byte heapsort updated shell to remove reference to gonnet_ removed all as its done by makefile now
added testing bit

\subsection{perfex}
\cc{Perfex} is a tool used to access performance counters on a Pentium 4. By default,
it prints out the number of cycles\footnote{The processor automatically
maintains this figure in a 40-bit register called P4\_TSC (Time Stamp Counter),
and perfex prints this value after running the program} which a program takes to
execute. This was used to create the graphs comparing the running time of sorts.
The gnuplot scripts for this were short and easily written.

Perfex was an alternative to using SimpleScalar. By using its hardware
registers, actual real world results could have been measured. Using perfex
instead of SimpleScalar was an option reviewed at the start, but the difficulty
involved in setting up a machine to use it was considerable. Additionally,
because perfex measures results on a real machine, only one type of cache and
branch predictor could have been measured. We have provided limited
measurements to supplement the SimpleScalar results.

\subsection{Software Predictor}
The software predictor used was a series of macros and functions which
maintained the state of a branch predictor at a particular point in the program.
Statistics about each branch were collected, and dumped into text files, from
which scripts created graphs to display them. Some sample code is shown in
Figure \ref{Sample code using the software predictor}.

@@

%TODO another sentence here
Each of the four $O(NlogN)$ sorts and radixsort, was treated in this fashion.

>>>>

\subsection{PapiEx}
\cc{PapiEx} is a tool used to access performance counters on a Pentium 4. It
uses the PAPI framework to count certain characteristics across a program's
execution. We used it to count the number of cycles a sort took to run. This
provided an accurate measure of how fast a sort executes on a Pentium 4. The
PapiEx homepage is at \n{http://icl.cs.utk.edu/~mucci/papiex/}.

PapiEx was used to create the graphs comparing the running time of sorts.
The gnuplot scripts for this were short and easily written by hand.

PapiEx was an alternative to using SimpleScalar. By using its hardware
registers, actual real world results could have been measured. Using PapiEx
instead of SimpleScalar was an option reviewed at the start, but because PapiEx
measures results on a real machine, only one type of cache and branch predictor
could have been measured. We have instead provided limited measurements to
supplement the SimpleScalar results.

\subsection{Software Predictor}
The software predictor used was a series of macros and functions which
maintained the state of a bimodal branch predictor at a particular point in the
program. Statistics about each branch were collected, and dumped into text
files, from which scripts created graphs to display them. Some sample code is
shown in Figure \ref{Sample code using the software predictor}.

@@

%TODO another sentence here
Each of the four $O(NlogN)$ sorts, radixsort and shellsort, was treated in this fashion.

\section{Testing details}
%TODO a lot of this may be in the first sorts chapter, maybe heapsort.
discuss: BIG + random.org, 64 bit in lamarca vs 32 bit, do_nothing and how it
affects cache hits, scaling in the software branch predictors. How to compare
and contrast varying results (branch vs softwre branhc; P4 vs simplescalar).
What factors are used for testing: /proc/cpuinfo output, cahce and branch
predictor data from P4, simplescalar inputs, software predictor, verfication of
sorts.
